### 2018-04-09 Manuscript_figures_2017_12_12.Rmd
Updated figure order and commented code

---
title: "Landscape heterogeneity simulation figures"
author: "Lauren White"
date: "April 3, 2018"
output: html_document
---

## Load packages
```{r}
library(doBy)
library(ggplot2)
library(viridis)
library(ggthemes)
library(scales)
library(lsr)
library(SDMTools)
library(cowplot)
```

## Load and update summary dataset
```{r}
setwd("C:/Users/law2y/OneDrive/R Code/IBM/August 2017")
merged_data<-read.csv("IBMsummary.csv")
merged_data<-merged_data[-which(merged_data$beta2==5),] #remove saturating beta_2 values

#Create character strings for landscape structure (Hurst exponent and proportion available habitat)
merged_data$Hurst<-NA
merged_data$proportion<-NA
merged_data$Hurst[which(merged_data$H==0.1)]<-"HP"
merged_data$Hurst[which(merged_data$H==0.5)]<-"MP"
merged_data$Hurst[which(merged_data$H==0.9)]<-"LP"

merged_data$proportion[which(merged_data$p==0.25)]<-"L"
merged_data$proportion[which(merged_data$p==0.5)]<-"M"
merged_data$proportion[which(merged_data$p==0.75)]<-"H"
merged_data$lstructure<-paste(merged_data$proportion, merged_data$Hurst, sep="/")

#Order factors
merged_data$betas<- factor(merged_data$betas, levels = c("0 0 0", "0 0 -1", "0 1 -1", "0 1 -0.5", "0 2 -1", "0 2 -0.5", "3 0 0", "3 0 -1", "3 1 -1", "3 1 -0.5", "3 2 -1", "3 2 -0.5", "6 0 0", "6 0 -1", "6 1 -1", "6 1 -0.5", "6 2 -1", "6 2 -0.5"))
merged_data$lstructure<- factor(merged_data$lstructure, levels = c("L/HP", "L/MP", "L/LP", "M/HP", "M/MP", "M/LP", "H/HP", "H/MP", "H/LP"))

#Set working directory of where to produce Figures
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Figures April 2018")
```

## Define multiplot function
```{r}
#' Multiplot function
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols), byrow=TRUE)
  }

  if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

##Figure 1
Random forest variable importance results
```{r}
library(randomForest)
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Random Forest Analysis")

#Load importance values (not scaled by SD)
imp_logit<-read.csv("imp_logit.csv")
imp_logitprev<-read.csv("logit_imp_maxprev.csv")
imp_logitdur<-read.csv("logit_imp_dur.csv")

logitSD<-read.csv("imp_logitSD.csv")
logitprevSD<-read.csv("imp_logitprevSD.csv")
logitdurSD<-read.csv("imp_logitdurSD.csv")

imp_logit$SD<-logitSD$x
imp_logitprev$SD<-logitprevSD$x
imp_logitdur$SD<-logitdurSD$x

#Order according to increasing node purity
imp_logit<-imp_logit[order(imp_logit$IncNodePurity),]
imp_logitdur<-imp_logitdur[order(imp_logitdur$IncNodePurity),] 
imp_logitprev<-imp_logitprev[order(imp_logitprev$IncNodePurity),]

#Remove infection probability (dummy variable--not varied in factorial design)
imp_logit<- imp_logit[!(imp_logit$X=="inf_prob"),]
imp_logitdur<- imp_logitdur[!(imp_logitdur$X=="inf_prob"),]
imp_logitprev<- imp_logitprev[!(imp_logitprev$X=="inf_prob"),]

#In ggplot
imp_logit<-imp_logit[order(imp_logit$X.IncMSE),] #order according to %increase in MSE

#imp_logit$SD<-rep(0.005, times=nrow(imp_logit))
A<-ggplot(imp_logit, aes(x=reorder(X, X.IncMSE), y=X.IncMSE)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black", # Use black outlines,
             size=.3) +      # Thinner lines
    # geom_errorbar(aes(ymin=X.IncMSE-SD, ymax=X.IncMSE+SD),
    #               size=.3,    # Thinner lines
    #               width=.2,
    #               position=position_dodge(.9)) +
    scale_x_discrete(element_blank(), labels = c("rec_rate" = expression(gamma),
    "beta1" =  expression(beta[1]),"beta2" = expression(beta[2]), "beta3" = expression(beta[3]),
    "percep" = "r", "density"="d")) +
     #ylab("Mean Decrease in Accuracy (%IncMSE)") +
    ylab("")+
    theme_bw()+
    ggtitle("(A)") + #" Variable Importance for Epidemic \nSuccess") +
    theme(axis.text=element_text(size=6), plot.title = element_text(size = 8)) +
    theme(plot.margin=unit(c(0.1,0.1,0,0), "cm"))+
    coord_flip()
    # theme_bw(axis.text=element_text(size=14),    axis.title=element_text(size=14,face="bold"), plot.title = element_text(size = 40))
A

#imp_logitprev$SD<-rep(0.0000001, times=nrow(imp_logitprev))
B<-ggplot(imp_logitprev, aes(x=reorder(X, X.IncMSE), y=X.IncMSE)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black", # Use black outlines,
             size=.3) +      # Thinner lines
    # geom_errorbar(aes(ymin=X.IncMSE-SD, ymax=X.IncMSE+SD),
    #               size=.3,    # Thinner lines
    #               width=.2,
    #               position=position_dodge(.9)) +
    scale_x_discrete(element_blank(), labels = c("rec_rate" = expression(gamma),
    "beta1" =  expression(beta[1]),"beta2" = expression(beta[2]), "beta3" = expression(beta[3]),
    "percep" = "r", "density"="d")) +
     #ylab("Mean Decrease in Accuracy (%IncMSE)") +
    ylab("")+
    theme_bw()+
    ggtitle("(B)") + #" Variable Importance for Maximum \nPrevalence|Success") +
    theme(axis.text=element_text(size=6), plot.title = element_text(size = 8)) +
    theme(plot.margin=unit(c(-0.25,0.1,0,0), "cm"))+
    coord_flip()
B

#imp_logitdur$SD<-rep(100, times=nrow(imp_logitdur))
C<-ggplot(imp_logitdur, aes(x=reorder(X, X.IncMSE), y=X.IncMSE)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black", # Use black outlines,
             size=.3) +      # Thinner lines
    # geom_errorbar(aes(ymin=X.IncMSE-SD, ymax=X.IncMSE+SD),
    #               size=.3,    # Thinner lines
    #               width=.2,
    #               position=position_dodge(.9)) +
    scale_x_discrete(element_blank(), labels = c("rec_rate" = expression(gamma),
    "beta1" =  expression(beta[1]),"beta2" = expression(beta[2]), "beta3" = expression(beta[3]),
    "percep" = "r", "density"="d")) +
    ylab("Mean Decrease in Accuracy (%IncMSE)") +
    theme_bw()+
    ggtitle("(C)") + #" Variable Importance for Epidemic \nDuration|Success") +

      theme(axis.text=element_text(size=6), plot.title = element_text(size = 8), axis.title.x=element_text(size=7)) +
    theme(plot.margin=unit(c(-0.25,0.1,0,0), "cm"))+
    coord_flip()
C

setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Figures April 2018")
tiff("Figure1_RF.tiff", height =10 , width =8.7, units = "cm", compression = "lzw", res = 1200)
multiplot(A, B, C, cols=1)
dev.off()

```

##Figure 1- vertical
Random forest variable importance results
```{r}
library(randomForest)
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Random Forest Analysis")

#Load importance values (not scaled by SD)
imp_logit<-read.csv("imp_logit.csv")
imp_logitprev<-read.csv("logit_imp_maxprev.csv")
imp_logitdur<-read.csv("logit_imp_dur.csv")

logitSD<-read.csv("imp_logitSD.csv")
logitprevSD<-read.csv("imp_logitprevSD.csv")
logitdurSD<-read.csv("imp_logitdurSD.csv")

imp_logit$SD<-logitSD$x
imp_logitprev$SD<-logitprevSD$x
imp_logitdur$SD<-logitdurSD$x

#Order according to increasing node purity
imp_logit<-imp_logit[order(imp_logit$IncNodePurity),]
imp_logitdur<-imp_logitdur[order(imp_logitdur$IncNodePurity),] 
imp_logitprev<-imp_logitprev[order(imp_logitprev$IncNodePurity),]

#Remove infection probability (dummy variable--not varied in factorial design)
imp_logit<- imp_logit[!(imp_logit$X=="inf_prob"),]
imp_logitdur<- imp_logitdur[!(imp_logitdur$X=="inf_prob"),]
imp_logitprev<- imp_logitprev[!(imp_logitprev$X=="inf_prob"),]

#In ggplot
imp_logit<-imp_logit[order(imp_logit$X.IncMSE),] #order according to %increase in MSE

#imp_logit$SD<-rep(0.005, times=nrow(imp_logit))
A<-ggplot(imp_logit, aes(x=reorder(X, X.IncMSE), y=X.IncMSE)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black", # Use black outlines,
             size=.3) +      # Thinner lines
    # geom_errorbar(aes(ymin=X.IncMSE-SD, ymax=X.IncMSE+SD),
    #               size=.3,    # Thinner lines
    #               width=.2,
    #               position=position_dodge(.9)) +
    scale_x_discrete(element_blank(), labels = c("rec_rate" = expression(gamma),
    "beta1" =  expression(beta[1]),"beta2" = expression(beta[2]), "beta3" = expression(beta[3]),
    "percep" = "r", "density"="d")) +
     ylab("Mean Decrease in Accuracy (%IncMSE)") +
    #ylab("")+
    theme_bw()+ theme(plot.margin = unit(c(0.1, 0, 0, 0), "cm"))+
    #ggtitle("(A)") + #" Variable Importance for Epidemic \nSuccess") +
    theme(axis.text=element_text(size=6), plot.title = element_text(size = 8), axis.title=element_text(size=6)) 
    #theme(plot.margin=unit(c(0.1,0.1,0,0), "cm"))
    # theme_bw(axis.text=element_text(size=14),    axis.title=element_text(size=14,face="bold"), plot.title = element_text(size = 40))
A

#imp_logitprev$SD<-rep(0.0000001, times=nrow(imp_logitprev))
B<-ggplot(imp_logitprev, aes(x=reorder(X, X.IncMSE), y=X.IncMSE)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black", # Use black outlines,
             size=.3) +      # Thinner lines
    # geom_errorbar(aes(ymin=X.IncMSE-SD, ymax=X.IncMSE+SD),
    #               size=.3,    # Thinner lines
    #               width=.2,
    #               position=position_dodge(.9)) +
    scale_x_discrete(element_blank(), labels = c("rec_rate" = expression(gamma),
    "beta1" =  expression(beta[1]),"beta2" = expression(beta[2]), "beta3" = expression(beta[3]),
    "percep" = "r", "density"="d")) +
     #ylab("Mean Decrease in Accuracy (%IncMSE)") +
    ylab(NULL)+
    theme_bw()+ theme(plot.margin = unit(c(0.25, 0, 0, 0), "cm"))+
    #ggtitle("(B)") + #" Variable Importance for Maximum \nPrevalence|Success") +
    theme(axis.text=element_text(size=6), plot.title = element_text(size = 8))
    #theme(plot.margin=unit(c(-0.25,0.1,0,0), "cm"))
    #coord_flip()
B

#imp_logitdur$SD<-rep(100, times=nrow(imp_logitdur))
C<-ggplot(imp_logitdur, aes(x=reorder(X, X.IncMSE), y=X.IncMSE)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black", # Use black outlines,
             size=.3) +      # Thinner lines
    # geom_errorbar(aes(ymin=X.IncMSE-SD, ymax=X.IncMSE+SD),
    #               size=.3,    # Thinner lines
    #               width=.2,
    #               position=position_dodge(.9)) +
    scale_x_discrete(element_blank(), labels = c("rec_rate" = expression(gamma),
    "beta1" =  expression(beta[1]),"beta2" = expression(beta[2]), "beta3" = expression(beta[3]),
    "percep" = "r", "density"="d")) +
    #ylab("Mean Decrease in Accuracy (%IncMSE)") +
    ylab(NULL)+
    theme_bw()+ theme(plot.margin = unit(c(0.1, 0, 0, 0), "cm"))+
    #ggtitle("(C)") + #" Variable Importance for Epidemic \nDuration|Success") +

      theme(axis.text=element_text(size=6), plot.title = element_text(size = 8), axis.title.x=element_text(size=7)) 
    #theme(plot.margin=unit(c(-0.25,0.1,0,0), "cm"))
    #coord_flip()
C

prow <- plot_grid(A, B, C,
           align = 'vh',
           labels = c("(A)", "(B)", "(C)"),
           hjust = -0.75,
           ncol = 3, label_size=7
           )
#paxis <- axis_canvas(prow, axis = "y") 
#ggdraw(insert_yaxis_grob(prow, paxis, grid::unit(.25, "null")))

setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Figures April 2018")
tiff("Figure1_RFvert.tiff", height =4.5, width =8.7, units = "cm", compression = "lzw", res = 1200)
#multiplot(A, B, C, cols=3)
prow
dev.off()

```

##Figure 2
Box plots rec rate vs. density
```{r}
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Figures April 2018")
sub_data<-merged_data[which(merged_data$max_I>1),] #successful outbreaks
colnames(sub_data)[3]<-"d" #rename density as "d"
colnames(sub_data)[11]<-"r" #rename perceptual range as "r"
#means <- aggregate(max_prevalence ~ density+ rec_rate +percep, sub_data, mean)
#means_dur <- aggregate(duration ~ density+ rec_rate +percep, sub_data, mean)

#tiff("boxplot_prev.tiff", height = 8, width = 12, units = "in", compression = "lzw", res = 300)
gg<- ggplot(data=sub_data, aes(y=max_prevalence, x=as.factor(rec_rate)))
#gg<- gg+ geom_bar(stat="identity", position=position_dodge(), colour="black")+ facet_grid(Rec.prob. ~ percep, scales= "free_x", labeller=label_both)
gg<-gg+ geom_boxplot()+ facet_grid(r ~ d, scales= "free_x", labeller=label_value)
gg<- gg+ labs(x = expression(Recovery~rate~(gamma)), y = "Maximum prevalence") #,title="(A)")
#gg<- gg+ theme(axis.title.x = element_text(face="bold", size=10),title=element_text(face="bold", size=10))
gg<- gg + theme(axis.text = element_text(size=6))
gg<- gg + theme(axis.text.x =element_text(vjust= 0, angle=-90))
gg<- gg + theme(axis.title = element_text(vjust=0.5, size=7))
gg<- gg + theme(strip.text = element_text(vjust=0.5, size=6))
gg<- gg + theme(plot.title = element_text(size = 7))
gg<- gg + theme(plot.margin=unit(c(0.1,0.1,0,0.1), "cm"))
#gg<-gg + geom_text(data = means , aes(label = duration, y = duration + 0.08))

gg
#dev.off()

#tiff("boxplot_dur.tiff", height = 8, width = 12, units = "in", compression = "lzw", res = 300)
hh<- ggplot(data=sub_data, aes(x=as.factor(rec_rate), y=duration))
#hh<- hh+ geom_bar(stat="identity", position=position_dodge(), colour="black")+ facet_grid(Rec.prob. ~ percep, scales= "free_x", labeller=label_both)
hh<-hh+ geom_boxplot()+ facet_grid(r ~ d, scales= "free_x", labeller=label_value)
hh<- hh+ labs(fill= "density", x = expression(Recovery~rate~(gamma)), y = "Duration") #, title="(B)")
#hh<- hh+ theme(axis.title = element_text(size=10), title=element_text(face="bold", size=10))
hh<- hh + theme(axis.text = element_text(vjust=0.5, size=6))
hh<- hh + theme(axis.text.x =element_text(vjust= 0, angle=-90))
hh<- hh + theme(axis.title = element_text(vjust=0.5, size=7))
hh<- hh + theme(plot.title = element_text(size=7)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=6)) 
hh<- hh + theme(plot.margin=unit(c(0.1,0.1,0,0.1), "cm"))
hh
#dev.off()

prow <- plot_grid(gg + theme(legend.position="none"),
            hh + theme(legend.position="none"),
           align = 'vh',
           labels = c("(A)", "(B)"),
           hjust = -1,
           ncol = 2, label_size=7
           )

tiff("Figure2.tiff", height = 7, width = 8.7, units = "cm", compression = "lzw", res = 1200)
prow
#multiplot(gg,hh, cols=2)
dev.off()

```

##Figure 3
```{r}
#percep=3, rec_rate=0.1, density=0.5
p3r0.1d0.5<-merged_data[which(merged_data$percep==3 & merged_data$rec_rate==0.1 & merged_data$density==0.5),]
ii<- ggplot(data=p3r0.1d0.5, aes(x=betas, y=max_prevalence,fill=as.factor(beta2)))
ii<- ii+ geom_boxplot()+ facet_grid( p~ H)
ii<- ii+ labs(x = expression(RSF~(beta[1]~beta[2]~beta[3])), y = "Maximum prevalence", fill= expression(beta[2]))
ii<- ii+ theme(plot.title = element_text(size=7))
ii<- ii+ theme(axis.title = element_text(size=6))
ii<- ii + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=6))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=6)) 
ii<- ii + theme(legend.text = element_text(vjust=0.5, size=6), legend.title=element_text(vjust=0.5, size=6))
ii<- ii + geom_vline(xintercept=c(6.5,12.5), linetype="dashed")
ii<- ii + theme(plot.margin=unit(c(0.1,0.1,0,0.1), "cm"))
ii

jj<- ggplot(data=p3r0.1d0.5, aes(x=betas, y=duration, fill=as.factor(beta2)))
jj<- jj+ geom_boxplot()+ facet_grid( p~ H)
jj<- jj+ labs(x = expression(RSF~(beta[1]~beta[2]~beta[3])), y = "Duration", fill= expression(beta[2]))
jj<- jj+ theme(plot.title = element_text(size=7))
jj<- jj+ theme(axis.title = element_text(size=6))
jj<- jj + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=6))
jj<- jj + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
jj<- jj + theme(strip.text = element_text(vjust=0.5, size=6)) 
jj<- jj + theme(legend.text = element_text(vjust=0.5, size=6), legend.title=element_text(vjust=0.5, size=7))
jj<- jj + geom_vline(xintercept=c(6.5,12.5), linetype="dashed")
jj<- jj + theme(plot.margin=unit(c(-0.1,0.1,0,0.1), "cm"))
jj

prow <- plot_grid( ii + theme(legend.position="none"),
            jj + theme(legend.position="none"),
           align = 'vh',
           labels = c("(A)", "(B)"),
           hjust = -1,
           nrow = 2, label_size=7
           )

legend_b <- get_legend(ii + theme(legend.position="bottom"))

p <- plot_grid( prow, legend_b, ncol = 1, rel_heights = c(1, .02))
p

tiff("Figure3test.tiff", height = 22.7, width = 17.8, units = "cm", compression = "lzw", res = 900)
#multiplot(ii,jj, cols=1)
p
dev.off()
```

##Figure 4
Comparing effects on landscape structure for select RSFs holding density, recovery rate and perceptual range constant.
```{r}
#sub_data<-merged_data[which(merged_data$max_I>1),]
sub_data<-merged_data

betas<-sub_data[which(sub_data$betas==c("3 0 0", "3 0 -1", "3 2 -0.5")),]
betas<-betas[which(betas$rec_rate==0.1 & betas$percep==3 & betas$density==0.5),]

#Individual plots of RSF box plot max prevalence and duration
ii<- ggplot(data=betas, aes(x=lstructure, y=max_prevalence)) #, fill=covariation)
ii<- ii+ geom_boxplot(stat="boxplot", position=position_dodge())
ii<- ii + facet_grid(betas ~.)
# ii<-ii +  scale_x_discrete(limits= c("L/HP", "L/MP", "L/LP", "M/HP", "M/MP", "M/LP", "H/HP", "H/MP", "H/LP"))
ii<- ii+ labs(x = "Landscape structure", y = "Maximum \n prevalence")
#ii<-ii +theme_grey()
#ii<-ii + theme(plot.title = element_text(hjust = 0.5))
ii<- ii+ theme(plot.title = element_text(size=7))
ii<- ii+ theme(axis.title = element_text(size=6))
ii<- ii + theme(axis.text.x = element_text(angle=90, hjust= 0, vjust=0.5, size=6))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=6)) 
#ii <- ii+ theme(legend.position="none")
#ii<-ii+  theme( axis.ticks.length=unit(-0.1, "cm")) #, axis.ticks.margin=unit(0.5, "cm"))
#ii<-ii+ theme(axis.ticks.x=element_blank())
#ii<-ii +theme (axis.text.x = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")), axis.text.y = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")))
ii

#Individual plots of RSF box plot max prevalence and duration
jj<- ggplot(data=betas, aes(x=lstructure, y=duration)) #, fill=covariation)
jj<- jj+ geom_boxplot(stat="boxplot", position=position_dodge())
jj<- jj + facet_grid(betas ~.)
# jj<-jj +  scale_x_discrete(limits= c("L/HP", "L/MP", "L/LP", "M/HP", "M/MP", "M/LP", "H/HP", "H/MP", "H/LP"))
jj<- jj+ labs(x = "Landscape structure", y = "Duration")
#jj<-jj +theme_grey()
#jj<-jj + theme(plot.title = element_text(hjust = 0.5))
jj<- jj+ theme(plot.title = element_text(size=7))
jj<- jj+ theme(axis.title = element_text(size=6))
jj<- jj + theme(axis.text.x = element_text(angle=90, hjust= 0, vjust=0.5, size=6))
jj<- jj + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
jj<- jj + theme(strip.text = element_text(vjust=0.5, size=6)) 
#jj <- jj+ theme(legend.position="none")
#jj<-jj+  theme( axis.ticks.length=unit(-0.1, "cm")) #, axis.ticks.margin=unit(0.5, "cm"))
#jj<-jj+ theme(axis.ticks.x=element_blank())
#jj<-jj +theme (axis.text.x = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")), axis.text.y = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")))
jj

prow <- plot_grid(ii, jj,
           align = 'vh',
           labels = c("(A)", "(B)"),
           hjust = -1,
           ncol = 1, label_size=7
           )

tiff("Figure4.tiff", height = 12, width = 8.7, units = "cm", compression = "lzw", res = 1200)
#multiplot(ii,jj, cols=1)
prow
dev.off()

```

##Supplementary Figure S1
Box plot of percep vs. rec rate vs. density for every RSF combination
```{r}
sub_data<-merged_data[which(merged_data$max_I>1),]
sub_data25<-sub_data[which(sub_data$density==0.25),]
sub_data50<-sub_data[which(sub_data$density==0.5),]

#tiff("boxplot_prev.tiff", height = 8, width = 12, units = "in", compression = "lzw", res = 300)
gg<- ggplot(data=sub_data25, aes(x=betas, y=max_prevalence))#, fill=as.factor(density)))
#gg<- gg+ geom_bar(stat="identity", position=position_dodge(), colour="black")+ facet_grid(Rec.prob. ~ percep, scales= "free_x", labeller=label_both)
gg<-gg+ geom_boxplot()+ facet_grid(rec_rate ~ percep, scales= "free_x", labeller=label_both)
gg<- gg+ labs(x = expression(Resource~Selection~Function~(beta[1]~beta[2]~beta[3])), y = "Maximum prevalence",title="(A)")
gg<- gg+ theme(axis.title = element_text(face="bold", size=20))
gg<- gg + theme(axis.text.x = element_text(angle=90, vjust=0.5, size=16))
gg<- gg + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
gg<- gg + theme(strip.text = element_text(vjust=0.5, size=14)) 
gg<- gg + theme(legend.text = element_text(vjust=0.5, size=16), legend.title=element_text(vjust=0.5, size=20))
gg
#dev.off()

#tiff("boxplot_dur.tiff", height = 8, width = 12, units = "in", compression = "lzw", res = 300)
hh<- ggplot(data=sub_data25, aes(x=betas, y=duration))#, fill=as.factor(density)))
#hh<- hh+ geom_bar(stat="identity", position=position_dodge(), colour="black")+ facet_grid(Rec.prob. ~ percep, scales= "free_x", labeller=label_both)
hh<-hh+ geom_boxplot()+ facet_grid(rec_rate ~ percep, scales= "free_x", labeller=label_both)
hh<- hh+ labs(x = expression(Resource~Selection~Function~(beta[1]~beta[2]~beta[3])), y = "Duration", title="(B)")
hh<- hh+ theme(axis.title = element_text(face="bold", size=20))
hh<- hh + theme(axis.text.x = element_text(angle=90, vjust=0.5, size=16))
hh<- hh + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=14)) 
hh<- hh + theme(legend.text = element_text(vjust=0.5, size=16), legend.title=element_text(vjust=0.5, size=20))
hh
#dev.off()

#tiff("boxplot_prev.tiff", height = 8, width = 12, units = "in", compression = "lzw", res = 300)
ii<- ggplot(data=sub_data50, aes(x=betas, y=max_prevalence))#, fill=as.factor(density)))
#gg<- gg+ geom_bar(stat="identity", position=position_dodge(), colour="black")+ facet_grid(Rec.prob. ~ percep, scales= "free_x", labeller=label_both)
ii<-ii+ geom_boxplot()+ facet_grid(rec_rate ~ percep, scales= "free_x", labeller=label_both)
ii<- ii+ labs(x = expression(Resource~Selection~Function~(beta[1]~beta[2]~beta[3])), y = "Maximum prevalence",title="(C)")
ii<- ii+ theme(axis.title = element_text(face="bold", size=20))
ii<- ii + theme(axis.text.x = element_text(angle=90, vjust=0.5, size=16))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=14)) 
ii<- ii + theme(legend.text = element_text(vjust=0.5, size=16), legend.title=element_text(vjust=0.5, size=20))
ii
#dev.off()

#tiff("boxplot_dur.tiff", height = 8, width = 12, units = "in", compression = "lzw", res = 300)
jj<- ggplot(data=sub_data50, aes(x=betas, y=duration))#, fill=as.factor(density)))
#jj<- jj+ geom_bar(stat="identity", position=position_dodge(), colour="black")+ facet_grid(Rec.prob. ~ percep, scales= "free_x", labeller=label_both)
jj<-jj+ geom_boxplot()+ facet_grid(rec_rate ~ percep, scales= "free_x", labeller=label_both)
jj<- jj+ labs(x = expression(Resource~Selection~Function~(beta[1]~beta[2]~beta[3])), y = "Duration", title="(D)")
jj<- jj+ theme(axis.title = element_text(face="bold", size=20))
jj<- jj + theme(axis.text.x = element_text(angle=90, vjust=0.5, size=16))
jj<- jj + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
jj<- jj + theme(strip.text = element_text(vjust=0.5, size=14)) 
jj<- jj + theme(legend.text = element_text(vjust=0.5, size=16), legend.title=element_text(vjust=0.5, size=20))
jj

tiff("FigureS2.tiff", height = 20, width = 14, units = "in", compression = "lzw", res = 300)
multiplot(gg,hh,ii,jj, cols=1)
dev.off()

```

##Supplementary Figure S2 (Heat maps, density=0.25)
Heatmap of max prevelance and mean duration for conspecific density=0.25 and comparing recovery rates of 0.1, 0.2, and 0.4
```{r}
r2<-merged_data[which(merged_data$rec_rate==0.2 & merged_data$density==0.25),]

sdf <- summaryBy(duration~ H + betas +p+ percep, data=r2, FUN=mean)
gg <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=duration.mean))
gg <- gg + geom_tile(color="white", size=0.1)
gg <- gg + scale_fill_viridis(name="Duration")
gg <- gg + coord_equal()
gg <- gg + facet_grid(percep ~ betas)
gg<- gg+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title= "(D)")
gg<- gg+ theme(axis.title = element_text(face="bold", size=24))
gg<- gg + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
gg<- gg + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
gg<- gg + theme(strip.text = element_text(vjust=0.5, size=18)) 
gg<- gg + theme(strip.text.x = element_text(angle = 90))
gg<- gg + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
gg

sdf <- summaryBy(max_prevalence~ H + betas +p+ percep, data=r2, FUN=mean)
hh <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=max_prevalence.mean))
hh <- hh + geom_tile(color="white", size=0.1)
hh <- hh + scale_fill_viridis(name="Mean \nmaximum \nprevalence")
hh <- hh + coord_equal()
hh <- hh + facet_grid(percep ~ betas)
hh<- hh+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title="(C)")
hh<- hh+ theme(axis.title = element_text(face="bold", size=24))
hh<- hh + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
hh<- hh + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=18)) 
hh<- hh + theme(strip.text.x = element_text(angle = 90))
hh<- hh + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
hh

r1<-merged_data[which(merged_data$rec_rate==0.1 & merged_data$density==0.25),]
sdf <- summaryBy(max_prevalence~ H + betas +p+ percep, data=r1, FUN=mean)
ii <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=max_prevalence.mean))
ii <- ii + geom_tile(color="white", size=0.1)
ii <- ii + scale_fill_viridis(name="Mean \nmaximum \nprevalence")
ii <- ii + coord_equal()
ii <- ii + facet_grid(percep ~ betas)
ii<- ii+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title="(A)")
ii<- ii+ theme(axis.title = element_text(face="bold", size=24))
ii<- ii + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=18)) 
ii<- ii + theme(strip.text.x = element_text(angle = 90))
ii<- ii + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
ii

sdf <- summaryBy(duration~ H + betas +p+ percep, data=r1, FUN=mean)
jj <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=duration.mean))
jj <- jj + geom_tile(color="white", size=0.1)
jj <- jj + scale_fill_viridis(name="Duration")
jj <- jj + coord_equal()
jj <- jj + facet_grid(percep ~ betas)
jj<- jj+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title="(B)")
jj<- jj+ theme(axis.title = element_text(face="bold", size=24))
jj<- jj + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
jj<- jj + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
jj<- jj + theme(strip.text = element_text(vjust=0.5, size=18)) 
jj<- jj + theme(strip.text.x = element_text(angle = 90))
jj<- jj + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
jj

r4<-merged_data[which(merged_data$rec_rate==0.4 & merged_data$density==0.25),]
sdf <- summaryBy(max_prevalence~ H + betas +p+ percep, data=r4, FUN=mean)
kk <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=max_prevalence.mean))
kk <- kk + geom_tile(color="white", size=0.1)
kk <- kk + scale_fill_viridis(name="Mean \nmaximum \nprevalence")
kk <- kk + coord_equal()
kk <- kk + facet_grid(percep ~ betas)
kk<- kk+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title="(E)")
kk<- kk+ theme(axis.title = element_text(face="bold", size=24))
kk<- kk + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
kk<- kk + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
kk<- kk + theme(strip.text = element_text(vjust=0.5, size=18)) 
kk<- kk + theme(strip.text.x = element_text(angle = 90))
kk<- kk + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
kk

sdf <- summaryBy(duration~ H + betas +p+ percep, data=r4, FUN=mean)
mm <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=duration.mean))
mm <- mm + geom_tile(color="white", size=0.1)
mm <- mm + scale_fill_viridis(name="Duration")
mm <- mm + coord_equal()
mm <- mm + facet_grid(percep ~ betas)
mm<- mm+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title="(F)")
mm<- mm+ theme(axis.title = element_text(face="bold", size=24))
mm<- mm + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
mm<- mm + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
mm<- mm + theme(strip.text = element_text(vjust=0.5, size=18)) 
mm<- mm + theme(strip.text.x = element_text(angle = 90))
mm<- mm + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
mm

tiff("FigureS4_heatmap.tiff", height = 27, width = 14, units = "in", compression = "lzw", res = 300)
multiplot(ii,jj,hh,gg, kk,mm, cols=1)
dev.off()
```

##Supplementary Figure S3 (Heat maps, density=0.5)
Heatmap of max prevelance and mean duration for conspecific density=0.5 and comparing recovery rates of 0.1, 0.2, and 0.4
```{r}
r2<-merged_data[which(merged_data$rec_rate==0.2 & merged_data$density==0.5),]

sdf <- summaryBy(duration~ H + betas +p+ percep, data=r2, FUN=mean)
gg <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=duration.mean))
gg <- gg + geom_tile(color="white", size=0.1)
gg <- gg + scale_fill_viridis(name="Duration")
gg <- gg + coord_equal()
gg <- gg + facet_grid(percep ~ betas)
gg<- gg+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title= "(D)")
gg<- gg+ theme(axis.title = element_text(face="bold", size=24))
gg<- gg + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
gg<- gg + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
gg<- gg + theme(strip.text = element_text(vjust=0.5, size=18)) 
gg<- gg + theme(strip.text.x = element_text(angle = 90))
gg<- gg + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
gg

sdf <- summaryBy(max_prevalence~ H + betas +p+ percep, data=r2, FUN=mean)
hh <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=max_prevalence.mean))
hh <- hh + geom_tile(color="white", size=0.1)
hh <- hh + scale_fill_viridis(name="Mean \nmaximum \nprevalence")
hh <- hh + coord_equal()
hh <- hh + facet_grid(percep ~ betas)
hh<- hh+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title="(C)")
hh<- hh+ theme(axis.title = element_text(face="bold", size=24))
hh<- hh + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
hh<- hh + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=18)) 
hh<- hh + theme(strip.text.x = element_text(angle = 90))
hh<- hh + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
hh

r1<-merged_data[which(merged_data$rec_rate==0.1 & merged_data$density==0.5),]
sdf <- summaryBy(max_prevalence~ H + betas +p+ percep, data=r1, FUN=mean)
ii <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=max_prevalence.mean))
ii <- ii + geom_tile(color="white", size=0.1)
ii <- ii + scale_fill_viridis(name="Mean \nmaximum \nprevalence")
ii <- ii + coord_equal()
ii <- ii + facet_grid(percep ~ betas)
ii<- ii+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title="(A)")
ii<- ii+ theme(axis.title = element_text(face="bold", size=24))
ii<- ii + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=18)) 
ii<- ii + theme(strip.text.x = element_text(angle = 90))
ii<- ii + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
ii

sdf <- summaryBy(duration~ H + betas +p+ percep, data=r1, FUN=mean)
jj <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=duration.mean))
jj <- jj + geom_tile(color="white", size=0.1)
jj <- jj + scale_fill_viridis(name="Duration")
jj <- jj + coord_equal()
jj <- jj + facet_grid(percep ~ betas)
jj<- jj+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title="(B)")
jj<- jj+ theme(axis.title = element_text(face="bold", size=24))
jj<- jj + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
jj<- jj + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
jj<- jj + theme(strip.text = element_text(vjust=0.5, size=18)) 
jj<- jj + theme(strip.text.x = element_text(angle = 90))
jj<- jj + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
jj

r4<-merged_data[which(merged_data$rec_rate==0.4 & merged_data$density==0.5),]
sdf <- summaryBy(max_prevalence~ H + betas +p+ percep, data=r4, FUN=mean)
kk <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=max_prevalence.mean))
kk <- kk + geom_tile(color="white", size=0.1)
kk <- kk + scale_fill_viridis(name="Mean \nmaximum \nprevalence")
kk <- kk + coord_equal()
kk <- kk + facet_grid(percep ~ betas)
kk<- kk+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title="(E)")
kk<- kk+ theme(axis.title = element_text(face="bold", size=24))
kk<- kk + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
kk<- kk + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
kk<- kk + theme(strip.text = element_text(vjust=0.5, size=18)) 
kk<- kk + theme(strip.text.x = element_text(angle = 90))
kk<- kk + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
kk

sdf <- summaryBy(duration~ H + betas +p+ percep, data=r4, FUN=mean)
mm <- ggplot(sdf, aes(x=as.factor(p), y=as.factor(H), fill=duration.mean))
mm <- mm + geom_tile(color="white", size=0.1)
mm <- mm + scale_fill_viridis(name="Duration")
mm <- mm + coord_equal()
mm <- mm + facet_grid(percep ~ betas)
mm<- mm+ labs(x="Proportion of available habitat (p)", y="Hurst exponent (H)", title="(F)")
mm<- mm+ theme(axis.title = element_text(face="bold", size=24))
mm<- mm + theme(axis.text.x = element_text(angle=90, hjust=0, vjust=0.5, size=16))
mm<- mm + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
mm<- mm + theme(strip.text = element_text(vjust=0.5, size=18)) 
mm<- mm + theme(strip.text.x = element_text(angle = 90))
mm<- mm + theme(legend.text = element_text(vjust=0.5, size=14), legend.title=element_text(vjust=0.5, size=20))
mm

tiff("FigureS3_heatmap.tiff", height = 27, width = 14, units = "in", compression = "lzw", res = 300)
multiplot(ii,jj,hh,gg, kk,mm, cols=1)
dev.off()
```

##Figure S4
```{r}
#percep=3, rec_rate=0.2, density=0.5
p3r0.2d0.5<-merged_data[which(merged_data$percep==3 & merged_data$rec_rate==0.2 & merged_data$density==0.5),]
ii<- ggplot(data=p3r0.2d0.5, aes(x=betas, y=max_prevalence, fill=as.factor(beta2)))
ii<- ii+ geom_boxplot()+ facet_grid( p~ H)
ii<- ii+ labs(x = expression(RSF~(beta[1]~beta[2]~beta[3])), y = "Maximum prevalence", title="(A)", fill= expression(beta[2]))
ii<- ii+ theme(plot.title = element_text(size=7))
ii<- ii+ theme(axis.title = element_text(size=6))
ii<- ii + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=6))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=6)) 
ii<- ii + theme(legend.text = element_text(vjust=0.5, size=6), legend.title=element_text(vjust=0.5, size=7))
ii<- ii + geom_vline(xintercept=c(6.5,12.5), linetype="dashed")
ii<- ii + theme(plot.margin=unit(c(0.1,0.1,0,0.1), "cm"))
ii

jj<- ggplot(data=p3r0.2d0.5, aes(x=betas, y=duration, fill=as.factor(beta2)))
jj<- jj+ geom_boxplot()+ facet_grid( p~ H)
jj<- jj+ labs(x = expression(RSF~(beta[1]~beta[2]~beta[3])), y = "Duration", title="(B)", fill= expression(beta[2]))
jj<- jj+ theme(plot.title = element_text(size=7))
jj<- jj+ theme(axis.title = element_text(size=6))
jj<- jj + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=6))
jj<- jj + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
jj<- jj + theme(strip.text = element_text(vjust=0.5, size=6)) 
jj<- jj + theme(legend.text = element_text(vjust=0.5, size=8), legend.title=element_text(vjust=0.5, size=7))
jj<- jj + geom_vline(xintercept=c(6.5,12.5), linetype="dashed")
jj<- jj + theme(plot.margin=unit(c(-0.1,0.1,0,0.1), "cm"))

jj

tiff("FigureS4.tiff", height = 22.5, width = 17.8, units = "cm", compression = "lzw", res = 900)
multiplot(ii,jj, cols=1)
dev.off()
```

##Figure S4 (log version)
```{r}
#percep=3, rec_rate=0.2, density=0.5
p3r0.2d0.5<-merged_data[which(merged_data$percep==3 & merged_data$rec_rate==0.2 & merged_data$density==0.5),]
ii<- ggplot(data=p3r0.2d0.5, aes(x=betas, y=log(max_prevalence), fill=as.factor(beta2)))
ii<- ii+ geom_boxplot()+ facet_grid( p~ H)
ii<- ii+ labs(x = expression(RSF~(beta[1]~beta[2]~beta[3])), y = "ln(Maximum prevalence)", title="(A)", fill= expression(beta[2]))
ii<- ii+ theme(plot.title = element_text(size=7))
ii<- ii+ theme(axis.title = element_text(size=6))
ii<- ii + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=6))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=6)) 
ii<- ii + theme(legend.text = element_text(vjust=0.5, size=6), legend.title=element_text(vjust=0.5, size=7))
ii<- ii + geom_vline(xintercept=c(6.5,12.5), linetype="dashed")
jj<- jj + theme(plot.margin=unit(c(0.1,0.1,0,0.1), "cm"))
ii

jj<- ggplot(data=p3r0.2d0.5, aes(x=betas, y=log(duration), fill=as.factor(beta2)))
jj<- jj+ geom_boxplot()+ facet_grid( p~ H)
jj<- jj+ labs(x = expression(RSF~(beta[1]~beta[2]~beta[3])), y = "ln(Duration)", title="(B)", fill= expression(beta[2]))
jj<- jj+ theme(plot.title = element_text(size=7))
jj<- jj+ theme(axis.title = element_text(size=6))
jj<- jj + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=6))
jj<- jj + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
jj<- jj + theme(strip.text = element_text(vjust=0.5, size=6)) 
jj<- jj + theme(legend.text = element_text(vjust=0.5, size=6), legend.title=element_text(vjust=0.5, size=7))
jj<- jj + geom_vline(xintercept=c(6.5,12.5), linetype="dashed")
jj<- jj + theme(plot.margin=unit(c(-0.1,0.1,0,0.1), "cm"))
jj

tiff("FigS4_log.tiff", height = 11.5, width = 14, units = "in", compression = "lzw", res = 300)
multiplot(ii,jj, cols=1)
dev.off()
```

##Fig. S5
Comparing effects on landscape structure for a single RSF(\beta_1=6, \beta_2=2, \beta_3=-0.5) for different densities, perceptual distances, and recovery rates.
```{r}
#sub_data<-merged_data[which(merged_data$max_I>1),]
sub_data<-merged_data

RSF620.5<-sub_data[which(sub_data$betas=="6 2 -0.5"),]
RSF620.5_rec0.1<-RSF620.5[which(RSF620.5$rec_rate==0.1),]
RSF620.5_rec0.2<-RSF620.5[which(RSF620.5$rec_rate==0.2),]
RSF620.5_rec0.4<-RSF620.5[which(RSF620.5$rec_rate==0.4),]

#Individual plots of RSF box plot max prevalence and duration
gg<- ggplot(data=RSF620.5_rec0.2, aes(x=lstructure, y=max_prevalence)) #, fill=covariation)
gg<- gg+ geom_boxplot(stat="boxplot", position=position_dodge())
gg<- gg + facet_grid(percep ~ density)
# gg<-gg +  scale_x_discrete(limits= c("L/HP", "L/MP", "L/LP", "M/HP", "M/MP", "M/LP", "H/HP", "H/MP", "H/LP"))
gg<- gg+ labs(title="(C)", x = "Landscape structure", y = "Mean maximum \n prevalence")
#gg<-gg +theme_grey()
#gg<-gg + theme(plot.title = element_text(hjust = 0.5))
gg<- gg+ theme(plot.title = element_text(size=7))
gg<- gg+ theme(axis.title = element_text(size=6))
gg<- gg + theme(axis.text.x = element_text(angle=90, hjust= 0, vjust=0.5, size=6))
gg<- gg + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
gg<- gg + theme(strip.text = element_text(vjust=0.5, size=6)) 
#gg <- gg+ theme(legend.position="none")
#gg<-gg+  theme( axis.ticks.length=unit(-0.1, "cm")) #, axis.ticks.margin=unit(0.5, "cm"))
#gg<-gg+ theme(axis.ticks.x=element_blank())
#gg<-gg +theme (axis.text.x = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")), axis.text.y = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")))
gg

hh<- ggplot(data=RSF620.5_rec0.2, aes(x=lstructure, y=duration)) #, fill=covariation)
hh<- hh+ geom_boxplot(stat="boxplot", position=position_dodge())
hh<- hh + facet_grid(percep ~ density)
# hh<-hh +  scale_x_discrete(limits= c("L/HP", "L/MP", "L/LP", "M/HP", "M/MP", "M/LP", "H/HP", "H/MP", "H/LP"))
hh<- hh+ labs(title="(D)", x = "Landscape structure", y = "Duration")
#hh<-hh +theme_grey()
#hh<-hh + theme(plot.title = element_text(hjust = 0.5))
hh<- hh+ theme(plot.title = element_text(size=7))
hh<- hh+ theme(axis.title = element_text(size=6))
hh<- hh + theme(axis.text.x = element_text(angle=90, hjust= 0, vjust=0.5, size=6))
hh<- hh + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=6)) 
#hh <- hh+ theme(legend.position="none")
#hh<-hh+  theme( axis.ticks.length=unit(-0.1, "cm")) #, axis.ticks.margin=unit(0.5, "cm"))
#hh<-hh+ theme(axis.ticks.x=element_blank())
#hh<-hh +theme (axis.text.x = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")), axis.text.y = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")))
hh

#Individual plots of RSF box plot max prevalence and duration
ii<- ggplot(data=RSF620.5_rec0.1, aes(x=lstructure, y=max_prevalence)) #, fill=covariation)
ii<- ii+ geom_boxplot(stat="boxplot", position=position_dodge())
ii<- ii + facet_grid(percep ~ density)
# ii<-ii +  scale_x_discrete(limits= c("L/HP", "L/MP", "L/LP", "M/HP", "M/MP", "M/LP", "H/HP", "H/MP", "H/LP"))
ii<- ii+ labs(x = "Landscape structure", y = "Mean maximum \n prevalence", title="(E)")
#ii<-ii +theme_grey()
#ii<-ii + theme(plot.title = element_text(hjust = 0.5))
ii<- ii+ theme(plot.title = element_text(size=7))
ii<- ii+ theme(axis.title = element_text(size=6))
ii<- ii + theme(axis.text.x = element_text(angle=90, hjust= 0, vjust=0.5, size=6))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=6)) 
#ii <- ii+ theme(legend.position="none")
#ii<-ii+  theme( axis.ticks.length=unit(-0.1, "cm")) #, axis.ticks.margin=unit(0.5, "cm"))
#ii<-ii+ theme(axis.ticks.x=element_blank())
#ii<-ii +theme (axis.text.x = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")), axis.text.y = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")))
ii

jj<- ggplot(data=RSF620.5_rec0.1, aes(x=lstructure, y=duration)) #, fill=covariation)
jj<- jj+ geom_boxplot(stat="boxplot", position=position_dodge())
jj<- jj + facet_grid(percep ~ density)
# jj<-jj +  scale_x_discrete(limits= c("L/HP", "L/MP", "L/LP", "M/HP", "M/MP", "M/LP", "H/HP", "H/MP", "H/LP"))
jj<- jj+ labs(x = "Landscape structure", y = "Duration", title="(F)")
#jj<-jj +theme_grey()
#jj<-jj + theme(plot.title = element_text(hjust = 0.5))
jj<- jj+ theme(plot.title = element_text(size=7))
jj<- jj+ theme(axis.title = element_text(size=6))
jj<- jj + theme(axis.text.x = element_text(angle=90, hjust= 0, vjust=0.5, size=6))
jj<- jj + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
jj<- jj + theme(strip.text = element_text(vjust=0.5, size=6)) 
#jj <- jj+ theme(legend.position="none")
#jj<-jj+  theme( axis.ticks.length=unit(-0.1, "cm")) #, axis.ticks.margin=unit(0.5, "cm"))
#jj<-jj+ theme(axis.ticks.x=element_blank())
#jj<-jj +theme (axis.text.x = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")), axis.text.y = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")))
jj

#Individual plots of RSF box plot max prevalence and duration
kk<- ggplot(data=RSF620.5_rec0.4, aes(x=lstructure, y=max_prevalence)) #, fill=covariation)
kk<- kk+ geom_boxplot(stat="boxplot", position=position_dodge())
kk<- kk + facet_grid(percep ~ density)
# kk<-kk +  scale_x_discrete(limits= c("L/HP", "L/MP", "L/LP", "M/HP", "M/MP", "M/LP", "H/HP", "H/MP", "H/LP"))
kk<- kk+ labs(title="(A)", x = "Landscape structure", y = "Mean maximum \n prevalence")
#kk<-kk +theme_grey()
#kk<-kk + theme(plot.title = element_text(hjust = 0.5))
kk<- kk+ theme(plot.title = element_text(size=7))
kk<- kk+ theme(axis.title = element_text(size=6))
kk<- kk + theme(axis.text.x = element_text(angle=90, hjust= 0, vjust=0.5, size=6))
kk<- kk + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
kk<- kk + theme(strip.text = element_text(vjust=0.5, size=6)) 
#kk <- kk+ theme(legend.position="none")
#kk<-kk+  theme( axis.ticks.length=unit(-0.1, "cm")) #, axis.ticks.margin=unit(0.5, "cm"))
#kk<-kk+ theme(axis.ticks.x=element_blank())
#kk<-kk +theme (axis.text.x = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")), axis.text.y = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")))
kk

mm<- ggplot(data=RSF620.5_rec0.4, aes(x=lstructure, y=duration)) #, fill=covariation)
mm<- mm+ geom_boxplot(stat="boxplot", position=position_dodge())
mm<- mm + facet_grid(percep ~ density)
# mm<-mm +  scale_x_discrete(limits= c("L/HP", "L/MP", "L/LP", "M/HP", "M/MP", "M/LP", "H/HP", "H/MP", "H/LP"))
mm<- mm+ labs(title="(B)", x = "Landscape structure", y = "Duration")
#mm<-mm +theme_grey()
#mm<-mm + theme(plot.title = element_text(hjust = 0.5))
mm<- mm+ theme(plot.title = element_text(size=7))
mm<- mm+ theme(axis.title = element_text(size=6))
mm<- mm + theme(axis.text.x = element_text(angle=90, hjust= 0, vjust=0.5, size=6))
mm<- mm + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
mm<- mm + theme(strip.text = element_text(vjust=0.5, size=6)) 
#mm <- mm+ theme(legend.position="none")
#mm<-mm+  theme( axis.ticks.length=unit(-0.1, "cm")) #, axis.ticks.margin=unit(0.5, "cm"))
#mm<-mm+ theme(axis.ticks.x=element_blank())
#mm<-mm +theme (axis.text.x = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")), axis.text.y = element_text(margin=unit(c(0.15,0.15,0.15,0.15), "cm")))
mm

tiff("FigureS5.tiff", height = 22, width = 17.8, units = "cm", compression = "lzw", res = 1200)
multiplot(kk,mm,gg,hh,ii,jj, cols=2)
dev.off()

```

##Supplementary Figure S6
See SimpleSIR.Rmd

##Supplementary Figure S7: Calibrating beta2/beta3 for RSF
```{r}
#Function to calculate rsf
myrsf=function(x,b1,b2) exp(x*b1+x^2*b2)   #caclulate rsf

#Function to calculate probability of selecting a given cell relative to other cells
#' @param x= \delta N
#' @param x.1= baseline population size of individuals
#' @param b2= \beta2 
#' @param b3= \beta3 (quadratic coefficient for RSF)
#' @param ncells= number of cells in neighborhood (e.g., 9, 25, 49)
myprsf=function(x,x.1,b2,b3,ncells)  #calculates rsf value for each cell; ncells= number of cells
{ lowlocs=myrsf(x.1,b2,b3)*(ncells-1)  #x = delta n
hiloc=myrsf(x+x.1,b2,b3)    
phi=hiloc/(hiloc+lowlocs)}

##beta2= 0, beta3=0
A <- ggplot(data.frame(x = c(0, 6)), aes(x = x)) +
  stat_function(fun = myprsf, args = list(0, 0, 0, 9),
                aes(linetype = "N=0")) +
  stat_function(fun =  myprsf, args = list(2, 0, 0, 9),
                aes(linetype = "N=2")) +
  stat_function(fun =  myprsf, args = list(4, 0, 0, 9),
                aes(linetype = "N=4")) +
  scale_x_continuous(name = expression(paste(Delta, "N")),
                     breaks = seq(0, 6, 1),
                     limits=c(0, 6)) +
  geom_hline(yintercept = 1/9, colour="red") + 
  scale_y_continuous(name = expression(paste("Probability of selecting cell with ", Delta, "N individuals")), limits=c(0,0.5)) +
  ggtitle(expression(paste("(A) ", beta[2], "= 0 & ", beta[3], "= 0 (Random)"))) +
  scale_linetype_manual("Initial \npopulation \nsize", values = c("solid", "dashed", "dotted"))
A

##beta2= 0, beta3=-1
B <- ggplot(data.frame(x = c(0, 6)), aes(x = x)) +
  stat_function(fun = myprsf, args = list(0, 0, -1, 9),
                aes(linetype = "N=0")) +
  stat_function(fun =  myprsf, args = list(2, 0, -1, 9),
                aes(linetype = "N=2")) +
  stat_function(fun =  myprsf, args = list(4, 0, -1, 9),
                aes(linetype = "N=4")) +
  scale_x_continuous(name = expression(paste(Delta, "N")),
                     breaks = seq(0, 6, 1),
                     limits=c(0, 6)) +
  geom_hline(yintercept = 1/9, colour="red") + 
  scale_y_continuous(name = expression(paste("Probability of selecting cell with ", Delta, "N individuals")), limits=c(0,0.5)) +
  ggtitle(expression(paste("(B) ", beta[2], "= 0 & ", beta[3], "= -1 (Negative)"))) +
  scale_linetype_manual("Initial \npopulation \nsize", values = c("solid", "dashed", "dotted"))
B

##beta2= 1, beta3=-1
C <- ggplot(data.frame(x = c(0, 6)), aes(x = x)) +
  stat_function(fun = myprsf, args = list(0, 1, -1, 9),
                aes(linetype = "N=0")) +
  stat_function(fun =  myprsf, args = list(2, 1, -1, 9),
                aes(linetype = "N=2")) +
  stat_function(fun =  myprsf, args = list(4, 1, -1, 9),
                aes(linetype = "N=4")) +
  scale_x_continuous(name = expression(paste(Delta, "N")),
                     breaks = seq(0, 6, 1),
                     limits=c(0, 6)) +
  geom_hline(yintercept = 1/9, colour="red") + 
  scale_y_continuous(name = expression(paste("Probability of selecting cell with ", Delta, "N individuals")), limits=c(0,0.5)) +
  ggtitle(expression(paste("(C) ", beta[2], "= 1 & ", beta[3], "=-1 (Positive)"))) +
  scale_linetype_manual("Initial \npopulation \nsize", values = c("solid", "dashed", "dotted"))
C

##beta2= 0, beta3=0
D <- ggplot(data.frame(x = c(0, 6)), aes(x = x)) +
  stat_function(fun = myprsf, args = list(0, 1, -0.5, 9),
                aes(linetype = "N=0")) +
  stat_function(fun =  myprsf, args = list(2, 1, -0.5, 9),
                aes(linetype = "N=2")) +
  stat_function(fun =  myprsf, args = list(4, 1, -0.5, 9),
                aes(linetype = "N=4")) +
  scale_x_continuous(name = expression(paste(Delta, "N")),
                     breaks = seq(0, 6, 1),
                     limits=c(0, 6)) +
  geom_hline(yintercept = 1/9, colour="red") + 
  scale_y_continuous(name = expression(paste("Probability of selecting cell with ", Delta, "N individuals")), limits=c(0,0.5)) +
  ggtitle(expression(paste("(D) ", beta[2], "= 1 & ", beta[3], "= -0.5 (Positive)"))) +
  scale_linetype_manual("Initial \npopulation \nsize", values = c("solid", "dashed", "dotted"))
D

##beta2= 2, beta3=-1
E <- ggplot(data.frame(x = c(0, 6)), aes(x = x)) +
  stat_function(fun = myprsf, args = list(0, 2, -1, 9),
                aes(linetype = "N=0")) +
  stat_function(fun =  myprsf, args = list(2, 2, -1, 9),
                aes(linetype = "N=2")) +
  stat_function(fun =  myprsf, args = list(4, 2, -1, 9),
                aes(linetype = "N=4")) +
  scale_x_continuous(name = expression(paste(Delta, "N")),
                     breaks = seq(0, 6, 1),
                     limits=c(0, 6)) +
  geom_hline(yintercept = 1/9, colour="red") + 
  scale_y_continuous(name = expression(paste("Probability of selecting cell with ", Delta, "N individuals")), limits=c(0,0.5)) +
  ggtitle(expression(paste("(E) ", beta[2], "= 2 & ", beta[3], "=-1 (Positive)"))) +
  scale_linetype_manual("Initial \npopulation \nsize", values = c("solid", "dashed", "dotted"))
E

##beta2= 2, beta3=-0.5
F <- ggplot(data.frame(x = c(0, 6)), aes(x = x)) +
  stat_function(fun = myprsf, args = list(0, 2, -0.5, 9),
                aes(linetype = "N=0")) +
  stat_function(fun =  myprsf, args = list(2, 2, -0.5, 9),
                aes(linetype = "N=2")) +
  stat_function(fun =  myprsf, args = list(4, 2, -0.5, 9),
                aes(linetype = "N=4")) +
  scale_x_continuous(name = expression(paste(Delta, "N")),
                     breaks = seq(0, 6, 1),
                     limits=c(0, 6)) +
  geom_hline(yintercept = 1/9, colour="red") + 
  scale_y_continuous(name = expression(paste("Probability of selecting cell with ", Delta, "N individuals")), limits=c(0,0.5)) +
  ggtitle(expression(paste("(F) ", beta[2], "= 2 & ", beta[3], "=-0.5 (Positive)"))) +
  scale_linetype_manual("Initial \npopulation \nsize", values = c("solid", "dashed", "dotted"))
F

tiff("RSF_conspecifics.tiff", height = 18, width = 14, units = "in", compression = "lzw", res = 300)
multiplot(A,B,C,D,E,F, cols=2)
dev.off()
```

##Figure S8
Representation of neutral landscapes used for simulation studies
```{r}
#' Create neutral landscape maps
library(raster)

#' Use standard methods to generate fractal maps. Binary and continuous surfaces may be produced.
#' 
#' @param k integer. The extent of the map (2^k+1)^2 pixels
#' @param h numeric. Level of aggregation in the map.
#' @param p numeric (0,1). The proportion of map in habitat=1
#' @param binary logical. If TRUE, a 0/1 categorical landscape is produced.
#' @author Shannon Pittman, James Forester, modified by Lauren White
#' @export
#' @example examples/neutral.landscape_example.R
fracland_mod <- function(k, h, p, binary = TRUE) {
  ## Function for creating neutral landscapes Shannon Pittman University of Minnesota May, 2013 k = the extent of the map (2^k+1)^2 pixels h =
  ## how clumped the map should be (ranging from ?? to ??) -- weird behavior at higher values p = proportion of map in habitat 1 binary =
  ## plotflag == if TRUE will plot a filled contour version of the matrix

  ## function call: testmap=land(6,1,.5,FALSE,TRUE)
  A <- 2^k + 1  # Scalar-determines length of landscape matrix

  #Right now, as written (1-p) represents the amount of habitat listed as "1"
  B <- matrix(0, A, A)  # Creates landscape matrix

  B[1, 1] <- 0
  B[1, A] <- 0
  B[A, 1] <- 0
  B[A, A] <- 0

  iter <- 1
  for (iter in 1:k) {
    scalef <- (0.5 + (1 - h)/2)^(iter)

    d <- 2^(k - iter)

    # ALL SQUARE STEPS#
    for (i in seq(d + 1, A - d, 2 * d)) {
      for (j in seq(d + 1, A - d, 2 * d)) {
        B[i, j] <- mean(c(B[i - d, j - d], B[i - d, j + d], B[i + d, j - d], B[i + d, j + d])) + scalef * rnorm(n = 1)
      }
    }

    # OUTSIDE DIAMOND STEP#
    for (j in seq(d + 1, A - d, 2 * d)) {
      B[1, j] <- mean(c(B[1, j - d], B[1, j + d], B[1 + d, j])) + scalef * rnorm(n = 1)
      B[A, j] <- mean(c(B[A, j - d], B[A, j + d], B[A - d, j])) + scalef * rnorm(n = 1)
    }

    for (i in seq(d + 1, A - d, 2 * d)) {
      B[i, 1] <- mean(c(B[i - d, 1], B[i + d, 1], B[i, 1 + d])) + scalef * rnorm(n = 1)
      B[i, A] <- mean(c(B[i - d, A], B[i + d, A], B[i, A - d])) + scalef * rnorm(n = 1)
    }

    # INSIDE DIAMOND STEP#
    if (2 * d + 1 <= A - 2 * d) {
      for (i in seq(d + 1, A - d, 2 * d)) {
        for (j in seq(2 * d + 1, A - 2 * d, 2 * d)) {
          B[i, j] <- mean(c(B[i - d, j], B[i + d, j], B[i, j - d], B[i, j + d])) + scalef * rnorm(n = 1)
        }
      }

      for (i in seq(2 * d + 1, A - 2 * d, 2 * d)) {
        for (j in seq(d + 1, A - d, 2 * d)) {
          B[i, j] <- mean(c(B[i - d, j], B[i + d, j], B[i, j - d], B[i, j + d])) + scalef * rnorm(n = 1)
        }
      }
    }

    iter <- iter + 1
  }

  if (binary == T) {
    R <- sort(B)
    PosR <- (1 - p) * length(R)  #larger values become habitat, designated as 1
    pval <- R[PosR]
    T1 <- which(B > pval)
    T2 <- which(B <= pval)
    B[T1] <- 1  #habitat is 1
    B[T2] <- 0
  } 
  return(B)
}

landscape<-fracland_mod(k=7, h=0.1, p=0.25)
map.1 <- raster(landscape)
one<-spplot(map.1)
landscape<-fracland_mod(k=7, h=0.1, p=0.5) 
map.2 <- raster(landscape)
two<-spplot(map.2)
landscape<-fracland_mod(k=7, h=0.1, p=0.75)
map.3 <- raster(landscape)
three<-spplot(map.3)

landscape<-fracland_mod(k=7, h=0.5, p=0.25) 
map.4 <- raster(landscape)
four<-spplot(map.4)
landscape<-fracland_mod(k=7, h=0.5, p=0.5) 
map.5 <- raster(landscape)
five<-spplot(map.5)
landscape<-fracland_mod(k=7, h=0.5, p=0.75)
map.6 <- raster(landscape)
six<-spplot(map.6)

landscape<-fracland_mod(k=7, h=0.9, p=0.25)
map.7 <- raster(landscape)
seven<-spplot(map.7)
landscape<-fracland_mod(k=7, h=0.9, p=0.5)
map.8 <- raster(landscape)
eight<-spplot(map.8)
landscape<-fracland_mod(k=7, h=0.9, p=0.75)
map.9 <- raster(landscape)
# nine<-spplot(map.9)

#convert the raster to points for plotting
H<-rep(0.1, times=length(map.1))
map.1 <- rasterToPoints(map.1)
p<-rep(0.25, times=length(map.1))
df <- data.frame(map.1, H, p)

map.2 <- rasterToPoints(map.2)
p<-rep(0.5, times=length(map.1))
df2 <- data.frame(map.2, H, p)

map.3 <- rasterToPoints(map.3)
p<-rep(0.75, times=length(map.1))
df3 <- data.frame(map.3, H, p)

H<-rep(0.5, times=length(map.1))
map.4 <- rasterToPoints(map.4)
p<-rep(0.25, times=length(map.1))
df4 <- data.frame(map.4, H, p)

map.5 <- rasterToPoints(map.5)
p<-rep(0.5, times=length(map.1))
df5 <- data.frame(map.5, H, p)

map.6 <- rasterToPoints(map.6)
p<-rep(0.75, times=length(map.1))
df6 <- data.frame(map.6, H, p)

H<-rep(0.9, times=length(map.1))
map.7 <- rasterToPoints(map.7)
p<-rep(0.25, times=length(map.1))
df7 <- data.frame(map.7, H, p)

map.8 <- rasterToPoints(map.8)
p<-rep(0.5, times=length(map.1))
df8 <- data.frame(map.8, H, p)

map.9 <- rasterToPoints(map.9)
p<-rep(0.75, times=length(map.1))
df9 <- data.frame(map.9, H, p)

#Make the points a dataframe for ggplot

df <- rbind(df, df2)
df <- rbind(df, df3)
df <- rbind(df, df4)
df <- rbind(df, df5)
df <- rbind(df, df6)
df <- rbind(df, df7)
df <- rbind(df, df8)
df <- rbind(df, df9)

my_x_title <- expression(paste("Hurst exponent- habitat patchiness (", italic("H"), ")"))
my_y_title <- expression(paste("Proportion of available habitat (", italic("p"), ")"))

#Now make the map
tiff("FigureS7_p_vs_H.tif", height = 8.7, width = 8.7, units = "cm", compression = "lzw", res = 900)
ggplot(data=df, aes(y=y, x=x)) +
  geom_raster(aes(fill=layer)) +
    theme_bw() +
  coord_equal() +
  facet_grid(p~ H) +
  #scale_fill_viridis(name="Resource quality")+
  #scale_fill_continuous(colours=c("#0000FFFF","#FFFFFFFF"))+
  scale_fill_gradient(low="lightgrey", high="darkgreen", guide=FALSE) +
  #scale_fill_continuous(guide = FALSE) +
  labs(y=my_y_title, x=my_x_title)+
  theme(axis.ticks = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank())+
  theme(axis.title = element_text(size=10)) +
  theme(strip.text = element_text(vjust=0.5, size=9))+ 
  theme(legend.text = element_text(vjust=0.5, size=9), legend.title=element_text(vjust=0.5, size=10))
dev.off()

```