### 2018-04-09 SimpleSIR.Rmd

---
title: "Simple SIR"
author: "Lauren White"
date: "February 24, 2018"
output: html_document
---

Deterministic and more simplistic stochastic SIR models for conspecific density vs. rec. rate combinations tested in spatially-explicit model. Assume homogenous mixing and no landscape structure.

#Load libraries
Load `deSolve` library for evaluating ordinary differntial equations (ODEs).
```{r setup}
library(deSolve)
library(ggplot2)
library(cowplot)
```

## Deterministic SIR Model
Code adapted from: <http://sbfnk.github.io/chennai-modelling/intro_to_r_gk.html>
```{r}
par(mfrow=c(1,1))
k<-8
dim<-2^k + 1
density<-0.25
n.initial <- round(density*dim^2) # inital population size

init <- c(S = n.initial-1, I = 1, R = 0)
times <- seq(0, 1000, by = 1) #1000 time steps

#frequency dependent transmission
freq_sir <- function(time, state, parameters) {
    with(as.list(c(state, parameters)), {
        dS = -beta * S * I/16512  # The change in S 
        dI = beta * S * I/16512 - gamma * I
        dR = gamma * I
        return(list(c(dS, dI, dR))) 
    })
}

#density dependent transmission function
density_sir <- function(time, state, parameters) {
    with(as.list(c(state, parameters)), {
        dS = -beta * S * I/257  # The change in S 
        dI = beta * S * I/257 - gamma * I
        dR = gamma * I
        return(list(c(dS, dI, dR))) 
    })
}

parameters <- c(beta = 0.2, gamma = 0.4)

out_freq <- ode(y = init, func = freq_sir, times = times, parms = parameters)
out_density <- ode(y = init, func = density_sir, times = times, parms = parameters)

plot(out_freq[, "time"], out_freq[, "I"]/n.initial, type = "l", lwd = 10, col = "green", xlab = "Time", 
    ylab = "Prevalence")
lines(out_density[, "time"], out_density[, "I"]/n.initial, col="red")
# lines(out[, "time"], out[, "R"], col = "blue", lwd = 10)
#legend("bottomright", c("Susceptibles", "Infecteds", "Recovereds"), lwd = 10, lty = 1, 
#    col = 2:4)  # Add legend and specific position

```

## Stochastic SIR Model 
Develop functions to run equivalent stochastic SIR model that assumes density-dependent transmission, homogenous mixing, and operates using a Reed Frost model.
```{r}
#' Function to randomly generate individuals and randomly assign 'nI' number of infected
#' individuals at t=0
#' @param n.initial- the number of desired individuals to start with
#' @param nI- the number of infected individuals to start with at t=0
#' @author Lauren White
make.inds <- function(n.initial, nI){
  id<-1:n.initial
  I<-sample(1:n.initial, nI)
  status<-rep("S", times=n.initial)
  status[I]<-"I"
  inds <- data.frame(id = id, status=status, stringsAsFactors=FALSE) 
  inds
}

#' Function that determines whether not a transmission event occurs
#' First checks to see if there are any individuals + infected infividuals present in the same cell
#' Then calculates final transmission probability based on # of infected individuals present
#' @param inds- dataframe with individual traits, including infection status
#' @param transProb- per interaction transmissioin probability
#' @author Lauren White
#' @date February 24, 2018
#' 
infection_simple<-function(inds, transProb){
    s_inds<-which(inds$status=="S" ) #which individuals are currently infected?
    if (length(s_inds>0)){
      i_count<-which(inds$status=="I" ) #which individuals are currently infected?
      finalProb <- 1 - (1 - transProb)^i_count
      transmit <- rbinom(length(s_inds), 1, finalProb)
      new_i<-s_inds[which(transmit == 1)]
      inds$status[new_i]<-"I"
  } 
  return(inds)
}

#############################################################################################
#############################################################################################

#' Function that allows infected individuals to recover at rate 'gamma'
#' @param inds: the dataframe describing the current status and location of individuals in the simulation
#' @param gamma: the recovery rate
#' @author Lauren White
recover.inds<-function(inds, gamma){
  infected<-which(inds$status=="I" ) #which individuals are currently infected?
  if(length(infected>0)){
    rec.prob<-runif(length(infected), min=0, max=1)
    for (i in 1:length(infected)){
      if(rec.prob[i]<= gamma){
        inds$status[infected[i]]<-"R"
      }
    }
  }
  return(inds)
}

SIR_simple_function<-function(maxtime, nsim, k, density, inf_prob, rec_rate){
  #Set up initial conditions for all simulations
  dim<-2^k + 1
  n.initial <- round(density*dim^2) # inital population size

  ##Initiate data structure across simulations
  mydata<-list(inds=NULL, N=NULL, infdat=NULL)
  filename<-paste("k", k, "dens", density, "inf", inf_prob, "rec", rec_rate, sep="")
  summary<-data.frame(k=rep(k, times=nsim), density=rep(density, times=nsim), inf_prob=rep(inf_prob, times=nsim), rec_rate=rep(rec_rate, times=nsim), duration=NaN, max_I=NaN, max_prevalence=NaN)
  infected<-matrix(NaN, nrow=maxtime, ncol=nsim)

  #Iterate for 'nsim' number of simulations per given parameter set
  for(count in 1:nsim){
    ##Initiate individuals
    inds <- make.inds(n.initial=n.initial, nI=1)
    mydata$inds[[count]]<-inds

    #Initiate longterm storage for infection data
    infdat<-inds$status

    # Object for storing results for single simulation (disease dynamics through time)
    N <-data.frame(S=NaN, I=NaN, R=NaN)
    N[1,] <- c(sum(inds$status=="S"), sum(inds$status=="I"), sum(inds$status=="R"))

    ##Run the simulation for a max of 'T' timesteps
    for(t in 2:maxtime){

      #Disease process(es)
      inds<-infection_simple(inds, transProb=inf_prob)
      inds<-recover.inds(inds, gamma=rec_rate)

      #Update disease status lists
      infdat<-cbind(infdat,inds$status)
      N[t,] <- c(sum(inds$status=="S"), sum(inds$status=="I"), sum(inds$status=="R"))
      if(sum(inds$status=="I")==0){ #If number of infected in individuals--> 0, stop running
        summary$duration[count]<-t-1
        for (i in t:maxtime){
          N[i,]<-N[t,] #automatically fill remaining time slots with current N values
        }
        break
      }
    }
    mydata$N[[count]]<-N
    mydata$infdat[[count]]<-infdat
    summary$max_I[count]<-max(N$I)
    summary$max_prevalence[count]<-max(N$I)/n.initial
    infected[,count]<-N$I
    print(count)
  }
  #Record output
  write.csv(summary,file=paste(filename,"summary.csv", sep="_"))
  write.csv(infected,file=paste(filename,"infected.csv", sep="_"))
}

```

##Run stochastic simulations
Run simplified SIR model for six combinations of conspecific density and recovery rate tested in spatially explicit model.

```{r}
# tic=Sys.time()
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/SimpleSIR")

#SIR_simple_function<-function(maxtime, nsim, k, density, inf_prob, rec_rate)
SIR_simple_function(1000, 100, 8, 0.25, 0.2, 0.1)
SIR_simple_function(1000, 100, 8, 0.25, 0.2, 0.2)
SIR_simple_function(1000, 100, 8, 0.25, 0.2, 0.4)
SIR_simple_function(1000, 100, 8, 0.5, 0.2, 0.1)
SIR_simple_function(1000, 100, 8, 0.5, 0.2, 0.2)
SIR_simple_function(1000, 100, 8, 0.5, 0.2, 0.4)

# print(difftime(Sys.time(),tic,units="mins"))

```

##Plot deterministic and stochastic SIR results (assuming equal mixing)
Compare/verify output from deterministic ODE and stochastic SIR models
```{r}
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/SimpleSIR")
#set wd to location of stochastic SIR simulation output files

par(mfrow=c(2, 3))
k<-8
dim<-2^k + 1
density<-0.25
n.initial25 <- round(density*dim^2) # inital population size

density<-0.5
n.initial50 <- round(density*dim^2) # inital population size

#Density=0.25
init <- c(S = n.initial25-1, I = 1, R = 0)
times <- seq(0, 100, by = 1) #1000 time steps

stoch<-read.csv("k8dens0.25inf0.2rec0.1_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.1))
head(out)

plot(out[, "time"], out[, "I"]/n.initial25, type = "l", lwd = 10, col = "green", xlab = "Time", 
    ylab = "Prevalence")
for(i in 1:(dim(stoch)[2]-1)){
lines(stoch$X, stoch[,i+1]/n.initial25, col = "red", lwd = 1)
}
title("density=0.25;beta=0.2;gamma=0.1")

stoch<-read.csv("k8dens0.25inf0.2rec0.2_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.2))
head(out)

plot(out[, "time"], out[, "I"]/n.initial25, type = "l", lwd = 10, col = "green", xlab = "Time", 
    ylab = "Prevalence")
for(i in 1:(dim(stoch)[2]-1)){
lines(stoch$X, stoch[,i+1]/n.initial25, col = "red", lwd = 1)
}
title("density=0.25;beta=0.2;gamma=0.2")

stoch<-read.csv("k8dens0.25inf0.2rec0.4_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.4))
head(out)

plot(out[, "time"], out[, "I"]/n.initial25, type = "l", lwd = 10, col = "green", xlab = "Time", 
    ylab = "Prevalence")
for(i in 1:(dim(stoch)[2]-1)){
  lines(stoch$X, stoch[,i+1]/n.initial25, col = "red", lwd = 1)
}
title("density=0.25;beta=0.2;gamma=0.4")

#Density=0.50
init <- c(S = n.initial50-1, I = 1, R = 0)
times <- seq(0, 100, by = 1) #1000 time steps

stoch<-read.csv("k8dens0.5inf0.2rec0.1_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.1))
head(out)

plot(out[, "time"], out[, "I"]/n.initial50, type = "l", lwd = 10, col = "green", xlab = "Time", 
    ylab = "Prevalence")
for(i in 1:(dim(stoch)[2]-1)){
lines(stoch$X, stoch[,i+1]/n.initial50, col = "red", lwd = 1)
}
title("density=0.5;beta=0.2;gamma=0.1")

stoch<-read.csv("k8dens0.5inf0.2rec0.2_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.2))
head(out)

plot(out[, "time"], out[, "I"]/n.initial50, type = "l", lwd = 10, col = "green", xlab = "Time", 
    ylab = "Prevalence")
for(i in 1:(dim(stoch)[2]-1)){
lines(stoch$X, stoch[,i+1]/n.initial50, col = "red", lwd = 1)
}
title("density=0.5;beta=0.2;gamma=0.2")

stoch<-read.csv("k8dens0.5inf0.2rec0.4_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.4))
head(out)

plot(out[, "time"], out[, "I"]/n.initial50, type = "l", lwd = 10, col = "green", xlab = "Time", 
    ylab = "Prevalence")
for(i in 1:(dim(stoch)[2]-1)){
  lines(stoch$X, stoch[,i+1]/n.initial50, col = "red", lwd = 1)
}
title("density=0.5;beta=0.2;gamma=0.4")
```

##Combine simulation ouput
```{r}
library(doBy)
library(ggplot2)
library(viridis)
library(ggthemes)
library(scales)
library(lsr)

##Read in file names 
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/SimpleSIR")
filenames <- list.files(path = "C:/Users/law2y/OneDrive/R Code/LandscapeCh/SimpleSIR")
summaries<-filenames[grep("summary", filenames)] #summary data
infecteds<-filenames[grep("infect", filenames)] #I data

##Make a merged dataset from all available .csv files
SIRdata_merged<- do.call("rbind", lapply(summaries, read.csv, header = TRUE))
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/")
write.csv(SIRdata_merged, file="SIRdata_merged.csv")

```

##Box plots
Box plots comparing density and recovery rates for homogenous mixing SIR model
```{r}
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Figures Dec. 2017")
#sub_data<-merged_data[which(merged_data$max_I>1),] #successful outbreaks

#tiff("boxplot_prev.tiff", height = 8, width = 12, units = "in", compression = "lzw", res = 300)
gg<- ggplot(data=SIRdata_merged, aes(y=max_prevalence, x=as.factor(rec_rate)))
#gg<- gg+ geom_bar(stat="identity", position=position_dodge(), colour="black")+ facet_grid(Rec.prob. ~ percep, scales= "free_x", labeller=label_both)
gg<-gg+ geom_boxplot()+ facet_grid(. ~ density, scales= "free_x", labeller=label_both)
gg<- gg+ labs(x = expression(Recovery~rate~(gamma)), y = "Maximum prevalence",title="(A)")
gg<- gg+ theme(axis.title = element_text(size=16),title=element_text(face="bold", size=16))
gg<- gg + theme(axis.text.x = element_text(vjust=0.5, size=16))
gg<- gg + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
gg<- gg + theme(strip.text = element_text(vjust=0.5, size=14)) 
gg
#dev.off()

means <- aggregate(duration ~ density+ rec_rate, SIRdata_merged, mean)
#tiff("boxplot_dur.tiff", height = 8, width = 12, units = "in", compression = "lzw", res = 300)
hh<- ggplot(data=SIRdata_merged, aes(x=as.factor(rec_rate), y=duration))
#hh<- hh+ geom_bar(stat="identity", position=position_dodge(), colour="black")+ facet_grid(Rec.prob. ~ percep, scales= "free_x", labeller=label_both)
hh<-hh+ geom_boxplot()+ facet_grid(. ~ density, scales= "free_x", labeller=label_both)
hh<- hh+ labs(fill= "density", x = expression(Recovery~rate~(gamma)), y = "Duration", title="(B)")
hh<- hh+ theme(axis.title = element_text(face="bold", size=16), title=element_text(face="bold", size=16))
hh<- hh + theme(axis.text.x = element_text(vjust=0.5, size=16))
hh<- hh + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=14))
hh<- hh + stat_summary(fun.y=mean, colour="darkred", geom="point", 
                           shape=18, size=3,show.legend = FALSE)
hh<-hh + geom_text(data = means , aes(label = duration, y = duration + 0.08))
hh
#dev.off()

tiff("DDtransmission.tif", height = 14, width = 8.7, units = "in", compression = "lzw", res = 300)
multiplot(gg,hh, cols=1)
dev.off()

```

##Fig. S6
Compare homogenous mixing and spatial models in box plot form.
```{r}
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/")
SIRdata_merged<-read.csv("SIRdata_merged.csv")
SIRdata_merged<-SIRdata_merged[,-1]

SIRdata_merged$percep<-"H"
SIRdata_merged$p<-NA
SIRdata_merged$H<-NA
SIRdata_merged$beta1<-NA
SIRdata_merged$beta2<-NA
SIRdata_merged$beta3<-NA
SIRdata_merged$betas<-NA
SIRdata_merged$pbyH<-NA
SIRdata_merged$Hurst<-NA
SIRdata_merged$proportion<-NA
SIRdata_merged$lstructure<-NA

h_vs_s <-rbind(merged_data, SIRdata_merged)
hh <- ggplot(h_vs_s, aes(x = as.factor(rec_rate), y = max_prevalence, fill = percep))
hh<-hh+ geom_boxplot()+ facet_grid(. ~ density, scales= "free_x", labeller=label_both)
hh<- hh+ labs(fill= "Perceptual \ndistance", x = expression(Recovery~rate~(gamma)), y = "Maximum prevalence")#, title="(A)")
hh<- hh+ theme(axis.title = element_text(size=6), plot.title=element_text(size=7))
hh<- hh + theme(axis.text.x = element_text(vjust=0.5, size=6))
hh<- hh + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=6))
hh<- hh + theme(legend.position="none")
hh<- hh + theme(legend.text = element_text(vjust=0.5, size=6), legend.title=element_text(vjust=0.5, size=7))
hh<- hh +theme(plot.margin=unit(c(0.1,0.1,0,0.1),"cm"))
#hh<- hh + stat_summary(fun.y=mean, colour="darkred", geom="point", 
                           #shape=18, size=3,show_guide = FALSE)
#hh<-hh + geom_text(data = means , aes(label = duration, y = duration + 0.08))
hh

ii <- ggplot(h_vs_s, aes(x = as.factor(rec_rate), y = duration, fill = percep))
ii<-ii+ geom_boxplot()+ facet_grid(. ~ density, scales= "free_x", labeller=label_both)
ii<- ii+ labs(fill= "Perceptual \ndistance", x = expression(Recovery~rate~(gamma)), y = "Duration") #, title="(B)")
ii<- ii+ theme(axis.title = element_text(size=6), plot.title=element_text(size=7))
ii<- ii + theme(axis.text.x = element_text(vjust=0.5, size=6))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=6))
ii<- ii + theme(legend.position="bottom")
ii<- ii + theme(legend.text = element_text(vjust=0.5, size=6), legend.title=element_text(vjust=0.5, size=7))
ii<- ii +theme(plot.margin=unit(c(-0.1,0.1,-0.5,0.1),"cm"))
#ii<- ii + stat_summary(fun.y=mean, colour="darkred", geom="point", 
                           #shape=18, size=3,show_guide = FALSE)
#ii<-ii + geom_text(data = means , aes(label = duration, y = duration + 0.08))
ii

#install.packages("ggpubr")
#library("ggpubr")
#ggarrange(hh, ii, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")

library(cowplot)
prow <- plot_grid( hh + theme(legend.position="none"),
            ii + theme(legend.position="none"),
           align = 'vh',
           labels = c("(A)", "(B)"),
           hjust = -1,
           nrow = 2, label_size=7
           )

# extract the legend from one of the plots
# (clearly the whole thing only makes sense if all plots
# have the same legend, so we can arbitrarily pick one.)
legend_b <- get_legend(ii + theme(legend.position="bottom"))

# add the legend underneath the row we made earlier. Give it 10% of the height
# of one plot (via rel_heights).
p <- plot_grid( prow, legend_b, ncol = 1, rel_heights = c(1, .07))
p

setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Figures April 2018")
tiff("FigureS6.tiff", height = 17, width = 17,8, units = "cm", compression = "lzw", res = 1200)
p
#multiplot(hh,ii, cols=1)
dev.off()
```